---
title: "AE 15: Ice cover in Madison, WI area lakes"
---

The data for this application exercise comes from the [lterdatasampler](https://lter.github.io/lterdatasampler/) package. The mission of the [Long Term Ecological Research program (LTER)](https://lternet.edu) Network is to "provide the scientific community, policy makers, and society with the knowledge and predictive understanding necessary to conserve, protect, and manage the nation’s ecosystems, their biodiversity, and the services they provide."

Specifically we'll be using data from the North Temperate Lakes LTER (NTL-LTER) site, which is located in the Madison, WI area, modeling the relationship between number of days that a lake is frozen, excluding periods where the lake thaws before refreezing again, and annual average temperature.

## Getting started

### Packages

We will use the **tidyverse** package for data visualization and wrangling and the **tidymodels** package for modeling.

```{r}
#| label: load-packages
#| message: false
#| warning: false
library(tidyverse)
library(tidymodels)
```

### Data

The data can be found in the `data` folder; it's called `icecover.csv`.

```{r}
#| label: load-data
#| message: false
icecover <- read_csv("data/icecover.csv")
```

The data dictionary is below:

| Variable Name      | Description                                                      |
|:-------------------|:-----------------------------------------------------------------|
| `lakeid`           | Lake name                                                        |
| `ice_on`           | Date of freeze of lake                                           |
| `ice_off`          | Date of ice breakup of lake                                      |
| `ice_duration`     | Number of days between the freeze and breakup dates of each lake |
| `year`             | Year of observation                                              |
| `annual_avg_temp`  | Annual average air temperature (°C)                              |

## Visualizing the model

We’re going to investigate the relationship between `ice_duration` and `annual_avg_temp`.

- Create an appropriate plot to investigate this relationship. Add appropriate labels to the plot.

```{r}
#| label: visualize
ggplot(icecover, aes(x = annual_avg_temp, y = ice_duration)) +
  geom_point() +
  labs(
    x = "Annual average temperature (°C)",
    y = "Ice duration (days)",
    title = "Ice duration vs. annual average temperature",
    subtitle = "North Temperate Lakes LTER"
  )
```

- If you were to draw a straight line to best represent the relationship between ice duration and annual average temperature, where would it go? Why?

- Now, let R draw the line for you. Refer to the documentation at <https://ggplot2.tidyverse.org/reference/geom_smooth.html>. Specifically, refer to the method section.

```{r}
#| label: visualize-with-line
ggplot(icecover, aes(x = annual_avg_temp, y = ice_duration)) +
  geom_point() +
  geom_smooth(method = "lm") +
  labs(
    x = "Annual average temperature (°C)",
    y = "Ice duration (days)",
    title = "Ice duration vs. annual average temperature",
    subtitle = "North Temperate Lakes LTER"
  )
```

- What types of questions can this plot help answer?

*How does ice duration change with annual average temperature? What is the predicted ice duration for a given annual average temperature?*

- We can use this line to make predictions. Predict what you think the ice duration would be in a year with annual average temperature of 7, 10, and 12 °C. Which prediction is considered extrapolation?

*At 7 °C, we estimate ice duration to be 100 days. At 10 °C, we estimate ice duration to be 80 days. 12 °C would be considered extrapolation because that is beyond the range of temperatures we actually observed in our data.*

## Model fitting

Fit a model to predict fish weights from their heights.

```{r}
#| label: fit-model
ice_temp_fit <- linear_reg() |>
  fit(ice_duration ~ annual_avg_temp, data = icecover)

ice_temp_fit
```

## Model summary

- Display the model summary including estimates for the slope and intercept along with measurements of uncertainty around them. Show how you can extract these values from the model output.

```{r}
#| label: model-summary
tidy(ice_temp_fit)
```

- Write out your model using mathematical notation.

$$
\widehat{ice~duration}=149.11-6.18\times temp
$$


## Prediction

Before, you used your eyeballs to predict what the ice duration would be in a year with annual average temperature of 7, 10, and 12 °C. Now let's have the computer do it. 

- First, use `R` as an overgrown calculator to compute the model prediction for each of the three temperatures by plugging into your model formulas and doing the arithmetic:

```{r}
#| label: predict-by-hand
148.109981 - 6.181554 * 7
148.109981 - 6.181554 * 10
148.109981 - 6.181554 * 12
```

- Second, use the `predict` function to perform the same calculation. You should get the same numbers (aside from rounding error).

```{r}
#| label: predict-by-r
new_lakes <- tibble(annual_avg_temp = c(7, 10, 12))

predict(ice_temp_fit, new_data = new_lakes)
```

## Correlation

We can also assess correlation between two quantitative variables.

- What is correlation? What are values correlation can take?

*The correlation coefficient is a number between -1 and 1 that measures the strength and direction of the linear association between two numerical variables.*

- What is the correlation between temperature and ice duration?

```{r}
#| label: correlation
icecover |>
  summarize(
    r = cor(annual_avg_temp, ice_duration)
  )
```

## Fitting other models

We can fit more models than just a straight line. Change the plotting code from earlier to use `method = "loess"`. What is different from the plot created before?

```{r}
#| label: visualize-with-curve
ggplot(icecover, aes(x = annual_avg_temp, y = ice_duration)) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(
    x = "Annual average temperature (°C)",
    y = "Ice duration (days)",
    title = "Ice duration vs. annual average temperature",
    subtitle = "North Temperate Lakes LTER"
  )
```
