---
title: "AE 23: Pokemon Inference"
---

Today we’ll explore the question “Are taller Pokémon speedier than shorter Pokémon?”

# Packages

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
library(tidymodels)
```

# Data

The data for this application exercise was originally gathered using the [**Complete Pokémon Dataset**](https://www.kaggle.com/datasets/rounakbanik/pokemon) from Kaggle. 

First, let's load the data:

```{r}
#| label: load-data
#| message: false

pokemon <- read_csv("data/pokemon.csv")


```

To keep things simple, we'll work with a subset of the data, the generation 4 Pokémon.

```{r}
pokemon <- pokemon |>
  filter(generation == 4)

glimpse(pokemon)
```

We are interested in whether a Pokémon's log height (measured in meters) can predict its speed (how fast it is).

# Visualize

a. Review! Create a new variable for the Pokémon's log height.

```{r}
pokemon <- pokemon |>
  mutate(log_height = log(height_m))
```


b. **Your turn**: Plot the data and the line of best fit.

```{r}
#| message: false
ggplot(pokemon, aes(x = log_height, y = speed)) + 
  geom_point() + 
  geom_smooth(method = "lm")
```

# Point Estimation

a.  **Your turn**: Fit the linear model to these data:

```{r}
observed_fit <- pokemon |>
  specify(speed ~ log_height) |>
  fit()
observed_fit
```

::: callout-note 
This gives the exact same numbers that you get if you use `linear_reg() |> fit()`, but we need this new syntax because it plays nice with the tools we have for confidence intervals and hypothesis tests. I know, I hate it too, but it's the way it is.
:::

b.  **Your turn**: Typeset the equation for the model fit:

$$
\widehat{\text{speed}} = 72.41 + 7.39\times \log(\text{height})
$$

c.  **Your turn**: Interpret the slope and the intercept estimates:

    -   If a Pokémon had 0 log(height) (equivalently 1 m tall), we predict that it would have a 72.41 speed stat on average;
    -   A \$1 increase in log(height) is associated with a 7.39 increase in speed, on average.


# Hypothesis Testing

Let's consider the hypotheses:

$$
H_0:\beta_1=0\quad vs\quad H_A: \beta_1\neq 0.
$$ The null hypothesis corresponds to the claim that log_height and speed are uncorrelated.

a.  Simulate and plot the null distribution for the slope:

```{r}
#| message: false
set.seed(24601)
null_dist <- pokemon |>
  specify(speed ~ log_height) |>
  hypothesize(null = "independence") |>
  generate(reps = 500, type = "permute") |>
  fit()

null_dist |>
  filter(term == "log_height") |>
  ggplot(aes(x = estimate)) + 
  geom_histogram()
```

b.  Add a vertical line to your plot indicating the point estimate of the slope from your original data analysis:

```{r}
visualize(null_dist) +
  shade_p_value(obs_stat = observed_fit, direction = "two-sided")
```

c.  Compute the $p$-value for this test and interpret it:

```{r}
#| warning: false
null_dist |>
  get_p_value(obs_stat = observed_fit, direction = "two-sided")
```

**Answer**: If the null were true and $\beta_1=0$ (meaning log_height and speed are uncorrelated), the probability of getting an estimate as or more extreme than the one we got ($b_1=7.39$) is 0.028. So we reject the null hypothesis. Evidence is sufficient to conclude that $\beta_1$, whatever it is, is not equal to zero.

# Interval Estimation

a.  **Demo**: Using seed `24601`, generate `500` bootstrap samples, and store them in a new data frame called `bstrap_samples`.

```{r}
set.seed(24601)
bstrap_samples <- pokemon |>
  specify(speed ~ log_height) |>
  generate(reps = 500, type = "bootstrap")
```

b.  **Demo**: Fit a linear model to each of these bootstrap samples and store the estimates in a new data framed called `bstrap_fits`.

```{r}
bstrap_fits <- bstrap_samples |> 
  fit()
```

c.  **Your turn**: Use `linear_reg() |> fit(...)` to fit a linear model to bootstrap sample number *347*, and verify that you get the same estimates as the ones contained in `bstrap_fits`.

```{r}
replicate_347 <- bstrap_samples |>
  filter(
    replicate == 347
  )

linear_reg() |>
  fit(speed ~ log_height, data = replicate_347) |>
  tidy()

bstrap_fits |>
  filter(replicate == 347)
```

::: callout-note 
The only point I'm making here is that this new bootstrap code is not performing a fundamentally new task. It's performing an old task (fitting the linear model), but it's repeating it A LOT. So the numbers you get here are not mysterious. They're numbers you already know how to compute.
:::

d.  **Demo**: Compute 95% confidence intervals for the slope and the intercept using the `get_confidence_interval` command.

```{r}
ci_95 <- get_confidence_interval(
  bstrap_fits,
  point_estimate = observed_fit,
  level = 0.95,
  type = "percentile"
)
ci_95
```

e.  **Your turn**: Verify that you get the same numbers when you manually calculate the quantiles of the slope estimates using `summarize` and `quantile`. *Pay attention to the grouping*.

```{r}
bstrap_fits |>
  ungroup() |>
  group_by(term) |>
  summarize(
    lower_ci = quantile(estimate, 0.025),
    upper_ci = quantile(estimate, 0.975)
  )
```

::: callout-note 
Same point as before. There's no magic here. `get_confidence_interval` is just a convenient way of doing something that you already knew how to do.
:::

f. **BONUS**: You can visualize the confidence interval:

```{r}
visualize(bstrap_fits) + 
  shade_confidence_interval(ci_95)
```


