---
title: Lab 8
subtitle: "Leavin' on a jet plane, confidently"
description: "Due: End of lab on Thurs, April 16"
categories: Lab
filters:
  - ../remove-fmt-skip.lua
draft: true
execute:
  echo: true
  eval: true
  message: false
editor: 
  markdown: 
    wrap: sentence
---

# Introduction

In this lab you'll explore the relationship between distance and air time of flights out of RDU in 2024.

And you'll work in what's considered a pretty unlikely situation: you have access to population data -- all flights out of RDU in 2024!
But we'll ask you to pretend that you didn't actually have access and you need to work with sample data.
How?
Hang in there, get the population data loaded first, then we'll explain the next steps!

## Getting started

By now you should be familiar with how to get started with a lab assignment by cloning the GitHub repo for the assignment.
If you're not sure how, refer back to an earlier lab.

Open the *lab-8.qmd* template Quarto file and update the `authors` field to add your name first (first and last) and then your teammates' names (first and last).
Render the document.
Examine the rendered document and make sure your and your teammates' names are updated in the document.
Commit and push your changes with a meaningful commit message and push to GitHub.

<details>

<summary>Click to expand if you need a refresher on assignment guidelines.</summary>

{{< include _guidelines.qmd >}}

</details>

## Packages

In this lab we will work with the **tidyverse** and **tidymodels** packages.

```{r}
#| eval: true
#| message: false
library(tidyverse)
library(tidymodels)
```

## Population data

The dataset, called `rdu-flights.csv`, can be found in the `data` folder.

# Questions

## Question 1 - Sample

Take a random sample of 100 flights and store it as `rdu_flights_sample`.
Make sure that you use the same seed as your teammates.

::: callout-tip
The following code allows you to obtain a random sample of N observations from a larger dataset.
Update the code below to complete Question 1.

```{r}
#| eval: false

set.seed(___)
df_sample <- df_full |>
  slice(sample(1:nrow(df_full), N))

```
:::

## Question 2 - Visualize

Visualize the relationship between `distance` (x) and `air_time` (y) using a scatter plot.
Add a regression line to the scatter plot, but do not display the standard error ribbon around the regression line. Comment on the observed relationship.

## Question 3 - Bootstrap

Compute a 95% bootstrap interval for the slope of the regression line for predicting `air_time` from `distance`.
In your code, use 1,000 bootstrap samples when simulating your bootstrap distribution. Note that you should have already set a seed in your answer to Question 1; thus, you should not set a new seed.

In your narrative, provide an interpretation of the 95% confidence interval for the slope (hint: refer back to "notes 22" on the course website).

::: callout-tip
Below is a step-by-step recipe for constructing and visualizing a confidence interval.
The code snippets shown are not "complete." They include some blanks you need to fill in, and they are just intended to guide you in the right direction.

-   Step 1: Calculate the point estimate for the slope and the intercept of the regression line.

```{r}
#| eval: false
obs_fit <- rdu_flights_sample |>
  specify(______ ~ ______) |>
  fit()
```

-   Step 2: Simulate a bootstrap distribution of regression estimates.

```{r}
#| eval: false
boot_dist <- rdu_flights_sample |>
  specify(_____ ~ ______) |>
  generate(reps = ____, type = _______) |>
  fit()
```

-   Step 3: Calculate the bounds of the confidence interval.

```{r}
#| eval: false
conf_ints <- 
  get_confidence_interval(
    ____, 
    level = ___, 
    point_estimate = ____
  )
```
:::

## Question 4: Hypothesize
a.  Set the null hypothesis:

    What if there was no relationship between `distance` and `air_time`? What would the true slope of the relationship be in that case? This is your null hypothesis. You can articulate it as:

    $H_0:$ There is no relationship between distance and air time, the value of the true slope is \_\_\_\_, i.e., $\beta_1$ = \_\_\_\_.

    Fill in the blanks.

    ::: callout-tip
    The null hypothesis **always** sets the true population parameter **equal** to a specific value.
    :::

b.  Set the alternative hypothesis:

    What if there is a relationship between `distance` and `air_time`? This is your alternative hypothesis!

    $H_A:$ There is a relationship between distance and air time, the value of the true slope is \_\_\_\_, i.e., $\beta_1$ \_\_\_\_\_\_.

    Fill in the blanks.

    ::: callout-tip
    The alternative hypothesis **always** compares the true population parameter to the same specific value set in the null hypothesis.
    This comparison can be "not equal to", "greater than", or "less than" and the choice depends on the research question.
    In this case, our research question is whether there is *some* relationship between distance and air time.
    :::

c. Based on the 95% confidence interval you came up with in Question 3, would you expect to fail to reject or to reject the above null hypothesis at the 5% significance level? Explain your reasoning.

# Wrap-up

::: callout-warning
Before you wrap up the assignment, make sure that you render, commit, and push one final time so that the final versions of both your `.qmd` file and the rendered PDF are pushed to GitHub and your Git pane is empty.
We will be checking these to make sure you have been practicing how to commit and push changes.
:::

## Submission

By now you should also be familiar with how to submit your assignment in Gradescope.

<details>

<summary>Click to expand if you need a refresher on how to get started with a lab assignment.</summary>

{{< include _submission.qmd >}}

</details>

## Grading and feedback

-   This lab is worth 30 points:
    -   10 points for being in lab and turning in *something* -- no partial credit for this part.
    -   20 points for:
        -   answering the questions correctly -- there is partial credit for this part.
        -   following the workflow -- there is partial credit for this part.
-   The workflow points are for:
    -   committing at least three times as you work through your lab,
    -   having your final version of `.qmd` and `.pdf` files in your GitHub repository, and
    -   overall organization.
-   You'll receive feedback on your lab on Gradescope within a week.

Good luck, and have fun with it!

